---
title: "Homework-03"
author: "Leah Pettway"
format: docx
---

Github repository: https://github.com/leahlamarre-ship-it/ENVS-193DS_homework-03

```{r}
#| label: set-up
#| message: false

library(tidyverse) #loading in packages
library(janitor) #loading in janitor
library(here) #loading in here
library(readxl) #loading in readxl

salinity <- read_csv("/Users/leahpettway/ENVS-193DS/github/ENVS-193DS_homework-03/HW3_data/salinity-pickleweed (1).csv") #loading in the data

my_data <- read_csv("/Users/leahpettway/ENVS-193DS/github/ENVS-193DS_homework-03/HW3_data/Personal_data.csv") #loading in personal data

```

## Part 2. Problems

### Problem 1. Slough soil salinity

#### a. An appropriate test

##### The appropriate tests to determine the strength of the relationship between salinity and California pickleweed biomnass is a pearson and spearman correlation test since both test how two variables are related to eachother or their strength of correlation. 

#### b. Create a visualization

```{r}
#| label: cleaning

salinity_clean <- salinity |> #new object created from salinity data frame
  clean_names() |> # cleaning up names so all are lowercase and no spaces
  rename("pickleweed_g" = "pickleweed") # including the unit pickle weed is in

```


```{r}
#| label: visualization of pickleweed and salinity

ggplot(data = salinity_clean, #use salinity_clean data frame
       aes(x = salinity_m_s_cm, #make the x-axis salinity_m_s_cm
           y = pickleweed_g)) + # make the y-axis pickleweed_g
  geom_point(color = "forestgreen") + # add point geometry
  labs(x = "Salinity (mS/cm)", #relabel the x-axis
       y = "Pickleweed biomass (g)") + # relabel the y-axis
  theme_light() #change theme from default

```

#### C. Checking Assumptions

##### Assumption of a linear relationship

```{r}
ggplot(data = salinity_clean, #use salinity_clean data frame
       aes(x = salinity_m_s_cm, #make the x-axis the predictor variable
           y = pickleweed_g)) + #make the y axis the response variable
  geom_point(color = "forestgreen") + #add point geometry to plot colored different from the default
  labs(x = "Salinity (mS/cm)", #relabel the x-axis
       y = "Pickleweed biomass (g)") + #relabel the y- axis
  theme_light() # change theme from default
```


##### Assumption of normal distribution

```{r}
#| label: QQ Plot

# base layer: ggplot call
ggplot(data = salinity_clean, # starting data frame using salinity_clean
       aes(sample = pickleweed_g)) + # y-axis for QQ plot 
  # first layer QQ reference line
  geom_qq_line(color = "red") + # using red so line is easier to see
  # second layer: QQ plot
  geom_qq()
  
```


```{r}
#| label: QQ Plot

# base layer: ggplot call
ggplot(data = salinity_clean, # starting data frame using salinity_clean
       aes(sample = salinity_m_s_cm)) + # y-axis for QQ plot 
  # first layer QQ reference line
  geom_qq_line(color = "red") + # using red so line is easier to see
  # second layer: QQ plot
  geom_qq() 

```


##### The assumptions I needed to check a few assumptions for a pearson correlation test, which  is a linear relationship between the variables and a normal distribution of of the variances for each variable. For the linear relationship I visually checked the data once I formatted it into a scatterplot to see the ralitonship between the two variables which appears to be linear due to the general trend of the points. I


















